Emergent dual scaling of riverine biodiversity

### 2021-08-17 manuscript_pnas.Rmd
Update manuscript_pnas.Rmd

**Significance** Larger ecosystems support more species; this ubiquitous pattern is the foundation of current conservation schemes. However, many ecosystems possess a complex spatial structure that cannot be represented by area, and the role of such complexity in regulating biodiversity patterns is largely unknown. Here, we use theory and extensive fish community data to show that ecosystem size and complexity dictate riverine biodiversity. We found that larger and more branched 'complex' river networks harbored greater species richness due to increased space and environmental heterogeneity. The complexity effect was comparable to the size effect, and this pattern has emerged regardless of ecological contexts. Our discovery illuminated a fundamental yet overlooked law in complex ecosystems, laying the groundwork for future research exploring macroecological patterns.

@@

Our unified framework unveiled the surprising consistency of dual scaling across ecological contexts or geographic regions, suggesting the universality of our findings. In particular, the emerging complexity-diversity relationship points to several important avenues for riverine biodiversity conservation. First and foremost, the simplification of complex river networks may result in regional biodiversity loss via reduced environmental diversity. For example, flow regulation [@poffHomogenizationRegionalRiver2007], habitat fragmentation [@grillMappingWorldFreeflowing2019], and stream burial [@elmoreDisappearingHeadwatersPatterns2008] may compromise or restrict access to the diverse habitats that complex branching networks may support. Hence, our results illuminate the importance of preserving the structural complexity of ecosystems, providing empirical support for.... Second, the complexity perspective may provide insights into the spatial planning of conservation efforts. Riverine reserves and local restoration are increasingly recognized as an effective management tool, and designing a spatial network of conservation sites in rivers is an area of active research [@koningNetworkGrassrootsReserves2020]. While large reserves or restoration sites are undoubtedly important, our results indicate the potential of coordinated networks of local conservation sites in protecting biodiversity. For example, synergies of multiple small reserves may emerge at the watershed scale when ecologically-distinct tributaries are involved in the design, as evidenced by the recent successful conservation of tropical fishes in Thailand's Salween basin [@koningNetworkGrassrootsReserves2020].

>>>>

**Significance** Larger ecosystems support more species; this ubiquitous pattern is the foundation of current conservation schemes. However, many ecosystems possess a complex spatial structure that cannot be represented by area, and the role of such complexity in regulating biodiversity patterns is largely unknown. Here, we use theory and extensive fish community data to show that ecosystem size and complexity dictate riverine biodiversity. We found that larger and more branched 'complex' river networks harbored greater species richness due to increased space and environmental heterogeneity. The complexity effect was comparable to the size effect, and this pattern has emerged regardless of ecological contexts. Our discovery illustrates a fundamental ecological law in complex ecosystems, laying the groundwork for future research exploring macroecological patterns.

@@

Our unified framework unveiled the surprising consistency of dual scaling across ecological contexts or geographic regions, suggesting the universality of our findings. In particular, the emerging complexity-diversity relationship points to several important avenues for riverine biodiversity conservation. First and foremost, any simplification of complex river networks may result in regional biodiversity loss via reduced environmental heterogeneity. Human-induced habitat alterations, including flow regulation [@poffHomogenizationRegionalRiver2007], habitat fragmentation [@grillMappingWorldFreeflowing2019], and stream burial [@elmoreDisappearingHeadwatersPatterns2008], may compromise or restrict access to the diverse habitats that complex branching networks may support. Hence, our results confirm the importance of preserving the structural complexity of ecosystems and support the modern conservation initiative that seeks to protect a diverse habitat portfolio [@schindlerPortfolioConceptEcology2015]. Second, the complexity perspective may provide insights into the spatial planning of conservation efforts. Riverine reserves and local restoration are increasingly recognized as an effective management tool, and the spatial prioritization of conservation sites in rivers is an area of active research [@koningNetworkGrassrootsReserves2020]. While the old "single large or several small" debate has led to a principle prioritizing a single large habitat [@diamondIslandDilemmaLessons1975], emerging perspectives posit the potential of coordinated networks of small habitats in protecting biodiversity [@fahrigWhySeveralSmall2020]. Understanding the role of branching complexity helps design such networks. For example, synergies of multiple small reserves may emerge at the watershed scale when ecologically-distinct tributaries are involved in the design, as evidenced by the recent successful conservation of tropical fishes in Thailand's Salween basin [@koningNetworkGrassrootsReserves2020].
